Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-12-26 paper/main.tex
add discussion of parameter estimation results

Figure~\ref{fig:parameter-estimates-single-step} shows profile-likelihoods for each optimized
parameter.
\textbf{TODO adapt these last sentences to new results}
We can see that all parameters can be estimated by our optimization approach and that no
non-identifiability~\cite{Raue2009} is present.
These plots also show that the individual growth rates of the bacteria can be estimated effectively.

>>>>

Figure~\ref{fig:parameter-estimates-single-step} shows profile-likelihoods for each optimized
parameter.
We compare four distinct models, covering Morse and Mie-type interactions with fixed damping
constant $\lambda=\SI{1}{\per\minute}$.
Subplot (A) displays the profile-likelihood for the damping constant $\lambda$ and exhibits a wide
range of permissible parameter values which motivates the approach to treat this parameter as a
constant thus reducing our overall model.
The parameters which characterize the Morse-type interaction are the potential stiffness (B),
strength (E) and radius (F).
While the potential stiffness (B) and radius (F) can be estimated well with the given data, the
potential strength (E) contains a practical unidentifiability~\cite{Raue2009} which is lifted when
fixing the damping constant $\lambda=\SI{1}{\per\minute}$.
In contrast to the Morse potential, the Mie potential uses two parameters $n,m$ (C,D) to determine
the shape of the interaction potential which allows it to describe a larger class of curves.
However, as subplots (A,C,D,E) show, neither the full Mie-type interaction nor a reduced model with
$\lambda=\SI{1}{\per\minute}$ allow for the estimation of the exponents $n,m$.
Although a minimum could be determined and the steepness of the enclosing profile towards the larger
values of exponents $n,m$ increases for the reduced model, the confidence intervals clearly show a
practical non-identifiability.
Since the reduction of the model improves the situation we are hopeful that an increased data set
could be suitable to estimate these parameters.
Notably, the two interaction types predict differing radii which are still within the $68\%$
confidence interval of each other.
The growth rates (G-L) can all be estimated for any interaction type but the non-reduced Mie-type
interaction shows wider confidence intervals, thus greater uncertainty in the predicted parameter
value.
